---
title: "College Majors"
author: "Stephen Wang"
date: "13/10/2019"
output: html_document
---

Data collected from: https://github.com/fivethirtyeight/data/tree/master/college-majors

When I was in highschool, I never gave my college major decision a thought. I thought going with the flow would be the best option which proved to be one of the biggest mistakes later down the road. So, although this is 5 years too late, the aim of this project is to explore data relating to the different college majors and find out which major will most likely ensure I have a job by the end of the day. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr) #data manipulation
library(tidyr) #tidying data 
library(readr) #reading files 
library(ggplot2) #data visualisation
library(purrr) #toolkit for functions and vectors
library(stringr) #string manipulation
library(reshape2) #reshaping data
library(ggcorrplot) #correlation matrixes
```

```{r}
filename = "recent-grads.csv"
grads <- read_csv(filename)
dim(grads)
head(grads, 10)
tail(grads, 10)
```

Dimension: 173 columns x 21 rows 

Variables:

Rank: rank by median earnings 
Major_code: unique major identifier 
Major: description of major
Total: total number of graduates with respective major
Men: total number of men gradutes with respective major
Women: total number of women graduates with respective major
Major_category: category of the major 
ShareWomen: proportion of women to total graduates with respective major
Sample_size:
Employed: number of employed graduates
Full_time: number of employed graduates with 35 hours or more
Part_time: number of employed graduates with less than 35 hours
Full_time_year_round:
Unemployed: number of unemployed graduates
Unemployment_rate: proportion of unemployed graduates (unemployed / unemployed+employed)
Median: median earnings of full time workers
P25th:
P75th:
College_jobs: number of jobs requiring degree
Non_college_jobs: number of jobs not requiring a degree
Low_wage_jobs: number of jobs that are low-wage service 

```{r}
colSums(is.na(grads))
subset(grads, is.na(Men) | is.na(Women))
```

```{r}
ggplot(data=grads,
       aes(x=reorder(Major_category, Major_category, 
                     function(x)-length(x))))+
  geom_bar(stat="count")+
  ylab("Number of Majors")+
  xlab("Category")+
  coord_flip()+
  theme_classic()
```

Engineering has more than double the second highest number of majors in a category. Let's take a closer look at the distribution of income in Science, Technology, Engineering, and Mathematics. 

Since the Rank variable is associated with Median, we can expect a clear negative relationship (as rank increases, median decreases). 

```{r}
science <- grads %>% filter(Major_category=='Health' | Major_category=='Biology & Life Science' | Major_category=='Psychology & Social Work') %>% mutate(stem='Science')

# technology <- grads %>% filter(Major_category=='Computers & Mathematics') %>% filter(grepl('COMPUTER|INFORMATION|COMMUNICATION', Major)) %>% mutate(stem='Technology')

technology <- grads %>% filter(Major_category=='Computers & Mathematics') %>% mutate(stem='Technology/Mathematics')

engineering <- grads %>% filter(Major_category=='Engineering') %>% mutate(stem='Engineering')

# mathematics <- grads %>% filter(Major_category=='Computers & Mathematics') %>% filter(!grepl('COMPUTER|INFORMATION|COMMUNICATION', Major)) %>% mutate(stem='Mathematics')

stem <- rbind(science, technology, engineering)

ggplot(data=stem,
       aes(x=stem, y=Median))+
  geom_boxplot()+
  theme_classic()+
  xlab(NULL)
```

The problem with dividing the data into these segments is that mathematics only have 3 sources of data while engineering have 29. I am making the assumption that the dataset combined Technology and Mathematics majors together on purpose so we'll be doing the same. 

```{r}
# Unemployment 
unemployment_desc <- grads[order(-grads$Unemployed),]
head(unemployment_desc, 5)

# Unemployment rate
unemployment_rate_desc <- grads[order(-grads$Unemployment_rate),]
head(unemployment_rate_desc, 5)


```

Note that the number of unemployed graduates are greater for popular major subjects but the unemployment rate is higher for niche major subjects i.e. nuclear engineering. 

Unemployment: when there are not enough opportunities to satisfy the number of graduates. 
